Preliminary Draft

### 2021-07-19 article/draft.tex
small rewriting

One of Facebook’s measures to regulate misinformation is to reduce the spread of misleading content through their ranking system. Facebook ranks each post and/or ad by assigning to it a relevancy score, where a high score leads to a high likelihood of the post and/or the ad to appear on a user's newsfeed. Doing so, Facebook can make a post or a whole account less visible by decreasing the relevancy score of its content; this is precisely the $reduce$ measure.\footnote{Lyons, T. (2018, May 22). The three-part recipe for cleaning up your news feed. Facebook Newsroom. \href{https://about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}{about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}.} This measure can be verified by looking at the number of views (reach) of a post, but this metric is not available via the CrowdTangle API or on Buzzsumo. Hence we can indirectly investigate the $reduce$ measure by looking at the engagement metrics (likes, comments, shares) related to a given post; which are available via the CrowdTangle API and Buzzsumo. If a post reaches less users because it has a lower ranking, then it is less likely to receive likes, comments and shares, relative to a post with a higher ranking.

@@

@@

The website $globalresearch.ca$ is linked to the Twitter account {$@CRG\_CRM$}; which was recently suspended.\footnote{We noticed the message about the account suspension on May 25, 2021. But to the best of our knowledge, no official communication by Twitter has announced the suspension nor the exact date at which it was implemented. Hence the account may have gotten suspended anytime between April 15, 2021 and May 25, 2021 (see the suspension screenshot in panel $(a)$ of figure \ref{fig3}). Furthermore, $globalresearch.ca$ has multiple failed fact-checks, see \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/}} When a user searches via the twitter search-box for any URL link of this website, no results appear as shown in the screenshot in panel $(b)$ of figure \ref{fig3}, taken on June $14$, $2021$. To further investigate the possible implementation of a reduced visibility measure, we search via the Twitter API for tweets, excluding retweets, containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $10$, $2021$.  As shown in panel (a) in figure \ref{fig4}, we find a strictly positive number of tweets containing the URL link {\it globalresearch.ca} throughout May $2021$ and the first week of June $2021$. Hence, the visibility of tweets containing this URL link has been reduced because users can no longer access tweets containing the URL link {\it globalresearch.ca} via the search box. Nevertheless users are not restrained from posting tweets containing this URL, as shown in the screenshot in panel $(a)$ of figure \ref{fig4bis}, found by taking the tweet ID of one of the collected tweets via the Twitter API. Furthermore, those collected Tweets have strictly positive engagement metrics as shown in panel $(b)$ of figure \ref{fig4}. Hence, the users who tweet articles from the {\it globalresearch.ca} website receive tweet level engagement from their own followers. Finally, when a user attempts to click on the URL link {\it globalresearch.ca} contained in the Tweet, a warning message appears and indicates that the link may be unsafe (see screenshot in panel (b) in figure \ref{fig4bis}).

@@

To the best of our knowledge, two types of flags can currently be seen on Facebook posts, videos and pictures: information banners that do not refute the message in the post but that give a link to an authoritative source (such as ‘Visit the COVID-19 Information Centre for vaccine resources’) and fact-check flags that provide a ‘judgement’ on the text of the post or the link shared (such as ‘False information Checked by independent fact-checkers.’) (see Figure \ref{fb_flags}). The fact-check flags can be of varied nature : False, Partly false, Missing context, False headline, Altered media, Opinion, Satire, Not eligible and even True.

@@

Surprinsingly only 11 posts out of the remaining 17 had a ‘False information’ flag  (see Table \ref{tab_flags_fb}). We observed that the flagged posts were also the ones in which the false link was expanded (i.e., a banner was visible with an image of the link and that can be clicked on, see the middle and right panels of Figure \ref{fb_flags} for examples). As the ‘False information’ flag is applied on the link banner, and not on the link itself, a user could share a link without the ‘False information’ banner as in the first example of Figure \ref{fb_flags}.

We also observed that most of the posts (13 out of 17, see Table \ref{tab_flags_fb}) had an information banner saying ‘Visit the COVID-19 Information Centre for vaccine resources’ (see the first two examples of Figure \ref{fb_flags}). We could not identify why the banner was applied on some posts and not on others (we saw no clear difference in the post messages for example). It should be noted that some posts displayed both the information and the fact-check banners, as in the right panel of Figure \ref{fb_flags}.

@@

Alongside other social networking platforms, when the content of a tweet violates the Twitter rules, a notice can be added to provide more context according to Twitter's Help Center.\footnote{See Notices on Twitter and what they mean: \href{https://help.twitter.com/en/rules-and-policies/notices-on-twitter}{https://help.twitter.com/en/rules-and-policies/notices-on-twitter}.} At the tweet level, notices take the form of a label or an interstitial. Labels are context specific (e.g. COVID19 or presidential elections) and  redirect users to a URL link to get more context (see figure \ref{fig8} for an example). Interstitials are presented as a greyed box on top of a tweet, which indicate sensitive content, violations of Twitter rules, withheld tweets for violation of local laws or even tweets from suspended accounts (see figure \ref{fig_notice} for an example). At the account level, notices can indicate whether an account has been temporarily or permanently suspended.

@@

To the best of our knowledge, when using the Twitter API v2, there is no field which indicates whether a tweet is labeled or not; while the interstitial ``possibly sensitive'' and ``withheld" are both Tweet fields that can be recovered\footnote{See \href{https://developer.twitter.com/en/docs/twitter-api/data-dictionary/object-model/tweet}{https://developer.twitter.com/en/docs/twitter-api/data-dictionary/object-model/tweet}.} from the API. Hence, in order to investigate the presence of labels, we resorted to scrape Twitter data using Minet [5]. This tool was recently enriched upon our request in order to capture whether a tweet contains a label or not, via the {\it minet twitter scrape} command.

@@

In this section, we take a deeper look at how labels and notices are introduced by Twitter, to indicate content which is inaccurate or false. To that end, we gathered a set of $3094$ URL links of articles which were marked as $False$ by Science Feedback, a fact-checking organization verifying the credibility of science-related viral information. As a second step, we collected (on June 30, 2021) via Minet Command line tool~\cite{minet} all the tweets that have shared a URL link which belongs to the set of $3094$ links marked as $False$. This data collection resulted in $323$ $938$ tweets, excluding retweets. Only $28$ tweets contained the label { ``Get the facts about COVID-19''}, $5$ tweets contained the label { ``Learn about US 2020 election security efforts''} and only $1$ had the following label {``This claim about election fraud is disputed''}. Furthermore, we noticed that the labeling rule might not be applied uniformly on a given set of tweets sharing the exact same URL link, among the set of collected tweets. More specifically, exactly $657$ tweets had shared a URL link redirecting to a video on Bitchute, entitled ``Important information on coronavirus 5G Kung Flu''. Among those $657$ tweets only $3$ contained the label ``Get the facts about COVID-19'' (see figure \ref{fig8}). This points towards the non-automation of the tweet labelling process and that it might be that only $3$ tweets got labelled after being reported by a user.

@@

\begin{figure}
	\centering
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_1.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_2.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_3.png}
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_4.png} 
	%\end{multicols}
	\caption{Top to buttom respectively: information panel displayed under some videos related to climate change, COVID19 vaccine, COVID19 and elections.}
	\label{fig9}
\end{figure}

@@

To further study the assignment of information panels to videos, we implemented the following exercice. We compiled a list of $171$ Youtube videos, checked by independent fact-checkers\footnote{Is it science Feedback ? if so add.} and marked as containing false or inaccurate information. We collected the information panels, when present, in each video by scrapping the content of the web and connecting to a US server. Four types of information panels were found in the list of visited videos as shown in figure\ref{fig9}:  information about COVID19, Vaccines, Climate change and US elections.

We classified the list of Youtube videos by topic based on its content as shown in table \ref{tab2}. We recorded the number of videos containing an information panel within each category. For COVID19 and vaccine related videos more than half of the videos contained an information panel. Nevertheless, we noticed that duplicates of the same Youtube video can get uploaded under different video titles and that the information panel appears under the some duplicates, but not for all. In addition, we noticed for COVID19 related videos, when the video title did not include keywords like (Testing, Pandemic, COVID, coronavirus), the video might not contain a panel associated with it, and in some cases when the video title contains variations of word COVID like (C.O.V.I.D or Cv19) it wouldn’t include a panel either. Therefore, we suspect that youtube is automatically adding panels to the videos based on the video title and not the content of the video since many of the videos that didn’t have a panel had the word COVID or coronavirus or vaccine mentioned in the video itself.

@@

For climate change we only examined $18$ misleading videos and we found that $28\%$ did not display an information panel. Lastly , for the general health category, we included the videos that were not related to COVID19 but contained misinformation concerning cures for cancer, abortion,  and viruses; no information panels were displayed for that set of Youtube videos. Therefore, it is likely that Youtube add information panels below videos concerning controversial topics that can have misleading information like the climate change, COVID19, flat earth and  vaccine. 
%drop last colum
%add % in a second line
\begin{table}[] 
\begin{tabular}{|c|c|c|c|c|}
\hline
Category       & \begin{tabular}[c]{@{}c@{}}Number of Youtube \\ videos with \\ an information panel\end{tabular} & \begin{tabular}[c]{@{}c@{}}Number of Youtube \\ videos without \\ an information panel\end{tabular}    \\ \hline
\begin{tabular}[c]{@{}c@{}} General \\ Health \end{tabular} & \begin{tabular}[c]{@{}c@{}} 0  \\ (0\%) \end{tabular}  & \begin{tabular}[c]{@{}c@{}} 11 \\ (100\%) \end{tabular}         \\ \hline
Vaccine        & \begin{tabular}[c]{@{}c@{}} 21     \\ (75\%) \end{tabular}   & \begin{tabular}[c]{@{}c@{}}  7 \\ (25\%)  \end{tabular}     \\ \hline
COVID19     & \begin{tabular}[c]{@{}c@{}} 73 \\ (63\%) \end{tabular}  & \begin{tabular}[c]{@{}c@{}} 43  \\ (37\%) \end{tabular}     \\ \hline
\begin{tabular}[c]{@{}c@{}}  Climate \\ change \end{tabular} & \begin{tabular}[c]{@{}c@{}} 5 \\ (28\%) \end{tabular}  & \begin{tabular}[c]{@{}c@{}} 13 \\ (72\%) \end{tabular}   \\ \hline
\end{tabular}
\caption{}
\label{tab2}
\end{table}

>>>>

One of Facebook’s measures to regulate misinformation is to reduce the spread of misleading content through their ranking system. Facebook ranks each post and/or ad by assigning to it a relevancy score, where a high score leads to a high likelihood of the post and/or the ad to appear on a user's newsfeed. Doing so, Facebook can make a post or a whole account less visible by decreasing the relevancy score of its content; this is precisely the $reduce$ measure.\footnote{Lyons, T. (2018, May 22). The three-part recipe for cleaning up your news feed. Facebook Newsroom. \href{https://about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}{about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}.} This measure can be verified by looking at the number of views (reach) of a post, but this metric is not available via the APIs used to access Facebook data: CrowdTangle or Buzzsumo. Hence we can indirectly investigate the $reduce$ measure by looking at the engagement metrics (likes, comments, shares) related to a given post; which are available on CrowdTangle and Buzzsumo. If a post reaches less users because it has a lower ranking, then it is less likely to receive likes, comments and shares, relative to a post with a higher ranking.

@@

We found irregularities in the number of Infowars articles collected from Buzzsumo. While Infowars usually publishes around 800 articles per month, only 53 articles were collected between June 11 to July 11, 2020\footnote{In Figure \ref{infowars2} the engagement metrics were only shown for days in which at least 5 Infowars articles were collected from Buzzsumo, to ensure reliable results.}. A temporary crawling problem coming from Buzzsumo may have caused this lack of data. Because no database is perfect, we would like to highlight the importance of cross-checking information between different sources when possible.

@@

The website $globalresearch.ca$ is linked to the Twitter account {$@CRG\_CRM$}; which was recently suspended.\footnote{We noticed the message about the account suspension on May 25, 2021. But to the best of our knowledge, no official communication by Twitter has announced the suspension nor the exact date at which it was implemented. Hence the account may have gotten suspended anytime between April 15, 2021 and May 25, 2021 (see the suspension screenshot in panel $(a)$ of figure \ref{fig3}).} When a user searches via the twitter search-box for any URL link of this website, no results appear as shown in the screenshot in panel $(b)$ of figure \ref{fig3}, taken on June $14$, $2021$.

To further investigate the possible implementation of a reduced visibility measure, we search via the Twitter API for tweets, excluding retweets, containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $10$, $2021$.  As shown in panel (a) in figure \ref{fig4}, we find a strictly positive number of tweets containing the URL link {\it globalresearch.ca} throughout May $2021$ and the first week of June $2021$. Hence, the visibility of tweets containing this URL link has been reduced because users can no longer access tweets containing the URL link {\it globalresearch.ca} via the search box. Nevertheless users are not restrained from posting tweets containing this URL, as shown in the screenshot in panel $(a)$ of figure \ref{fig4bis}, found by taking the tweet ID of one of the collected tweets via the Twitter API. Furthermore, those collected Tweets have strictly positive engagement metrics as shown in panel $(b)$ of figure \ref{fig4}. Hence, the users who tweet articles from the {\it globalresearch.ca} website receive tweet level engagement only from their own followers.

Finally, when a user attempts to click on the URL link {\it globalresearch.ca} contained in the Tweet, a warning message appears and indicates that the link may be unsafe (see screenshot in panel (b) in figure \ref{fig4bis}).

@@

To the best of our knowledge, two types of flags can currently be seen on Facebook posts, videos and pictures: information banners that do not refute the message in the post but that give a link to an authoritative source (such as ‘Visit the COVID-19 Information Centre for vaccine resources’) and fact-check flags that provide a ‘judgement’ on the text of the post or the link shared (such as ‘False information Checked by independent fact-checkers.’) (see Figure \ref{fb_flags}). The fact-check flags can be of varied nature : `False', `Partly false', `Missing context', `False headline', `Altered media', `Opinion', `Satire', `Not eligible' and even `True'.

@@

Surprinsingly only 11 posts out of the remaining 17 had a ‘False information’ flag  (Table \ref{tab_flags_fb}, see left panel of Figure \ref{fb_flags} for an example). We observed that the flagged posts were also the ones in which the false link was expanded (i.e., a banner was visible with an image of the link and that can be clicked on, see the middle and right panels of Figure \ref{fb_flags} for examples). As the ‘False information’ flag is applied on the link banner, and not on the link itself, a user is thus able to share a False link on Facebook without the ‘False information’ banner if the link is not expanded.

We also observed that most of the posts sharing this False link (13 out of 17, Table \ref{tab_flags_fb}) had an information banner saying ‘Visit the COVID-19 Information Centre for vaccine resources’ (see the first two examples of Figure \ref{fb_flags}). We could not identify why the banner was applied on some posts and not on others (we saw no clear difference in the post messages for example). It should be noted that some posts displayed both the information and the fact-check banners, as in the middle panel of Figure \ref{fb_flags}.

@@

Alongside other social networking platforms, when the content of a tweet violates the Twitter rules, a notice can be added to provide more context according to Twitter's Help Center.\footnote{See Notices on Twitter and what they mean: \href{https://help.twitter.com/en/rules-and-policies/notices-on-twitter}{https://help.twitter.com/en/rules-and-policies/notices-on-twitter}.} At the tweet level, notices take the form of a label or an interstitial. Labels are context specific (e.g. COVID19 or presidential elections) and  redirect users to a URL link to get more context (for example `Get the facts about COVID-19', see right panel of figure \ref{fig8}). Interstitials are presented as a greyed box on top of a tweet, which indicate sensitive content, violations of Twitter rules, withheld tweets for violation of local laws or even tweets from suspended accounts (for example `The following media includes potentially sensitive content', see left panel of figure \ref{fig_notice}). At the account level, notices can also indicate whether an account has been temporarily or permanently suspended.

@@

To the best of our knowledge, when using the Twitter API v2, there is no field which indicates whether a tweet is labeled or not; while the interstitial ``possibly sensitive'' and ``withheld" are both Tweet fields that can be recovered\footnote{See \href{https://developer.twitter.com/en/docs/twitter-api/data-dictionary/object-model/tweet}{https://developer.twitter.com/en/docs/twitter-api/data-dictionary/object-model/tweet}.} from the API. Hence, in order to investigate the presence of labels of all types, we resorted to scrape Twitter data using Minet [5]. This tool was recently enriched upon our request in order to capture whether a tweet contains a label or not, via the {\it minet twitter scrape} command.

@@

In this section, we take a deeper look at how labels and notices are introduced by Twitter, to indicate content which is inaccurate or false. To that end, we gathered a set of $3094$ URL links of articles which were marked as $False$ by Science Feedback, a fact-checking organization verifying the credibility of science-related viral information. As a second step, we collected (on June 30, 2021) via Minet Command line tool~\cite{minet} all the tweets that have shared a URL link which belongs to the set of $3094$ links marked as $False$. This data collection resulted in $323$ $938$ tweets, excluding retweets. Only $28$ tweets contained the label { ``Get the facts about COVID-19''}, $5$ tweets contained the label { ``Learn about US 2020 election security efforts''} and only $1$ had the following label {``This claim about election fraud is disputed''}. Furthermore, we noticed that the labeling rule might not be applied uniformly on a given set of tweets sharing the exact same URL link, among the set of collected tweets. More specifically, exactly $657$ tweets had shared a URL link redirecting to a video on Bitchute, entitled ``Important information on coronavirus 5G Kung Flu''. Among those $657$ tweets only $3$ contained the label ``Get the facts about COVID-19'' (see figure \ref{fig8}). This points towards the non-automation of the tweet labelling process and that it might be that these $3$ tweets were the only ones reported by a user.

@@

\begin{figure}
	\centering
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_1.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_2.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_3.png}
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_4.png} 
	%\end{multicols}
	\caption{Examples of YouTube information panels displayed under some videos.}
	\label{fig9}
\end{figure}

@@

To further study the assignment of information panels to videos, we compiled a list of $171$ Youtube videos, checked by independent fact-checkers\footnote{Is it science Feedback ? if so add.} and marked as containing false or inaccurate information. We collected the information panels, when present, in each video by connecting to a US server and scraping the content of the web page. Four types of information panels were found in the list of visited videos as shown in figure \ref{fig9}:  information about COVID19, Vaccines, Climate change and US elections.

We classified the list of Youtube videos by topic based on its content as shown in table \ref{tab2}. We recorded the number of videos containing an information panel within each category. For COVID19 and vaccine related videos more than half of the videos contained an information panel. Nevertheless, we noticed that duplicates of the same Youtube video can get uploaded under different video titles and that the information panel appears under the some duplicates, but not for all. In addition, we noticed for COVID19 related videos, when the video title did not include keywords like (Testing, Pandemic, COVID, coronavirus), the video might not contain a panel associated with it, and in some cases when the video title contains variations of word COVID like (C.O.V.I.D or Cv19) it wouldn’t include a panel either. Therefore, we suspect that youtube is automatically adding panels to the videos based on the video title and not the content of the video.

@@

For climate change we found that most of the videos (13 out of 18) did not display an information panel. Lastly, no information panels were found in the general health category set, that included videos that were not related to COVID19 but contained misinformation concerning cures for cancer, abortion, and viruses. Therefore, it is likely that Youtube add information panels below videos concerning controversial topics that can have misleading information like the climate change, COVID19, flat earth and vaccine. 
%drop last colum
%add % in a second line
\begin{table}[] 
\begin{tabular}{|c|c|c|c|c|}
\hline
Category       & \begin{tabular}[c]{@{}c@{}}Number of Youtube \\ videos with \\ an information panel\end{tabular} & \begin{tabular}[c]{@{}c@{}}Number of Youtube \\ videos without \\ an information panel\end{tabular}    \\ \hline
\begin{tabular}[c]{@{}c@{}} General \\ Health \end{tabular} & \begin{tabular}[c]{@{}c@{}} 0  \\ (0\%) \end{tabular}  & \begin{tabular}[c]{@{}c@{}} 11 \\ (100\%) \end{tabular}         \\ \hline
Vaccine        & \begin{tabular}[c]{@{}c@{}} 21     \\ (75\%) \end{tabular}   & \begin{tabular}[c]{@{}c@{}}  7 \\ (25\%)  \end{tabular}     \\ \hline
COVID19     & \begin{tabular}[c]{@{}c@{}} 73 \\ (63\%) \end{tabular}  & \begin{tabular}[c]{@{}c@{}} 43  \\ (37\%) \end{tabular}     \\ \hline
\begin{tabular}[c]{@{}c@{}}  Climate \\ change \end{tabular} & \begin{tabular}[c]{@{}c@{}} 5 \\ (28\%) \end{tabular}  & \begin{tabular}[c]{@{}c@{}} 13 \\ (72\%) \end{tabular}   \\ \hline
\end{tabular}
\caption{}
\label{tab2}
\end{table}
